---
title: "A fast algorithm for S-estimators"
author: "Matias Salibian"
date: "`r format(Sys.Date())`"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A fast algorithm for S-estimators for linear regression

This repository contains stand-alone `R` code implementing
the fast algorithm for S-regression estimators
proposed in 
[Salibian-Barrera, M. and Yohai, V.J. (2006).](http://dx.doi.org/10.1198/106186006X113629). 
The algorithm is also implemented 
in the function `robustbase::lmrob` that 
computes S- and MM-estimators. See the R package [robustbase](http://cran.r-project.org/web/packages/robustbase/index.html)
for more information.

Here is a simple example on how to use this code. We first 
read the function:
```{r initial}
source('fasts.R')
```
We will apply it to the well-known Boston Housing data set. 
```{r datasetup}
data(Boston, package='MASS')
x <- model.matrix(medv ~ ., data=Boston)
y <- Boston$medv
sest <- fast.s(x=x, y=y, int=0, N=500, k=2, best.r=5, seed=123)
```
The arguments of `fast.s` are:

* x: design matrix, covariates for each observation in rows
* y: vector of responses
* int: if equal to 1 a column of ones is added to x. This corresponds to including an intercept in the fitted model.
* N:  number of random sub-samples to be generated
* k: number of refining iterations in each subsample. k = 0 means "raw-subsampling"
* b: right hand side of the M-scale equation. Defaults to 0.50 yielding a highly-robust
estimator. 
* cc: tuning constant corresponding to the argument `b` above to ensure consistency.
* best.r: number of partially refined candidates to retain. These are finally iterated until convergence.
* seed: seed used in the generation of the random sub-samples. 

We can compare the S- and LS-estimators:
```{r compare}
lmest <- lm(medv ~ ., data=Boston)
round(cbind(sest[[1]], coef(lmest)), 3)
```
The estimated residual scales are
```{r compareses}
round(c(sest[[2]], summary(lmest)$sigma), 2)
```

I will later add here an illustration of the better predictions obtained
with the robust estimator.


